An Efficient Calculus for Generalised Parikh Images of Regular Languages

### 2022-04-05 paper/main.tex
Minor fudge changes

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  The image of the Parikh map is important in automata theory, offering a
  compact characterisation of an automaton. We contribute a novel understanding
  of how Parikh maps can be combined with arbitrary commutative monoid morphisms
  to efficiently represent a wide range of logics on automata and automata-like
  structures. This allows us to \Fudge{ensure world peace forever and express
  loop invariants that solve the halting problem for any practical program}.
  Furthermore, we show how this formulation can be efficiently implemented as a
  calculus in a theorem prover supporting Presburger logic. In particular, our
  calculus allows us to solve the common problem in string solvers of computing
  the Parikh image of products of arbitrarily many automata. We demonstrate this
  by implementing a tool called \Catra{}, which we use to solve a set of
  constraints produced by \Fudge{the PyEx benchmarks} when solved by the
  \Ostrich{} string constraint solver. We show that this implementation in
  addition to \Fudge{having a lower memory footprint} than the standard eager
  approach executed on the same underlying solver also \Fudge{strictly}
  outperforms the nuxmv model checker on the same problem\Fudge{, as well as the
  over-approximation recently described by~\citeauthor{approximate-parikh} and
  the original approach described by \citeauthor{generate-parikh-image}.}

@@

\Catra{} is implemented in Scala, with the calculus described in this paper
implemented as a theory plug-in for the \Princess{} automated theorem
prover~\cite{princess}. For comparison, we also provide an implementation of the
baseline method from~\cite{generate-parikh-image}, a direct translation that
uses the~\Nuxmv{} symbolic model checker~\cite{nuxmv} to solve our constraints,
and \Fudge{the approximation described in~\cite{approximate-parikh}}. An example
of an input file \Fudge{illustrating every feature of the file format} can be
seen in Listing~\ref{lst:input-example}.

@@

In standard mode, supported by all back-ends, \Catra{} tries to satisfy the
constraints expressed by the input file, reporting satisfiable or unsatisfiable
much like traditional~SAT- or SMT solvers. Additionally, \Fudge{the baseline and} our own back-end also supports generating the Presburger formula describing the constraints of the input file \Fudge{using the method described in Section~\ref{}}

@@

\contents{
  \item Future extensions: more logics
  \item Loop invariants?
  \item Integration into string solvers?
}

>>>>

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  The image of the Parikh map is important in automata theory, offering a
  compact characterisation of an automaton. We contribute \Fudge{a novel}
  understanding of how Parikh maps can be combined with arbitrary commutative
  monoid morphisms to efficiently represent a wide range of logics on automata
  and automata-like structures. This allows us to \Fudge{ensure world peace
  forever and express loop invariants that solve the halting problem for any
  practical program}. Furthermore, we show how this formulation can be
  efficiently implemented as a calculus in a theorem prover supporting
  Presburger logic. In particular, our calculus allows us to solve the common
  problem in string solvers of computing the Parikh image of products of
  arbitrarily many automata. We demonstrate this by implementing a tool called
  \Catra{}, which we use to solve a set of constraints produced by \Fudge{the
  PyEx benchmarks} when solved by the \Ostrich{} string constraint solver. We
  show that this implementation in addition to \Fudge{having a lower memory
  footprint} than the standard eager approach executed on the same underlying
  solver also \Fudge{strictly} outperforms the \Nuxmv{} model checker on the
  same problem\Fudge{, as well as both the over-approximation recently described
  by~\citeauthor{approximate-parikh} and the original approach described by
  \citeauthor{generate-parikh-image}.}

@@

\Catra{} is written in Scala, with the calculus described in this paper
implemented as a theory plug-in for the \Princess{} automated theorem
prover~\cite{princess}, which also performs the Presburger reasoning. For
comparison, we also provide an implementation of the baseline method
from~\cite{generate-parikh-image}, a direct translation that uses the~\Nuxmv{}
symbolic model checker~\cite{nuxmv} to solve our constraints, and \Fudge{the
approximation described in~\cite{approximate-parikh} on top of the standard
Verma back-end}. An example of an input file \Fudge{illustrating every feature
of the file format} can be seen in Listing~\ref{lst:input-example}.

@@

In satisfaction mode, supported by all back-ends, \Catra{} tries to satisfy the
constraints expressed by the input file, reporting satisfiable with register
assignments or unsatisfiable much like traditional~SAT- or SMT solvers would.
Additionally, \Fudge{the baseline and} our own back-end also supports generating
the Presburger formula describing the constraints of the input file \Fudge{using
the method described in Section~\ref{sec:presburger-representation}}

@@

\contents{
  \item Future extensions: more logics
  \item Loop invariants?
  \item Integration into string solvers?
  \item We have shown fastest
  \item We have shown versatile
  \item We have shown concretely useful
}
